---
title: HPI Data Analysis
author: by MZM
output: github_document
---

## Summary
In our project we aim to show how different countries compare against one another in multiple variables, one of which being the 'Happy Planet Index'. This will help us understand which regions or countries are able to provide a good living standard while also having a healthy environmental footprint. Our data is from a data set named the 'Happy Planet Index' which has taken most countries and used their social happiness, environmental impact and also inequalities and differences of the countries people. Some important variables to know are the 'Happy Planet Index' which is a combination of a countries inequalities of living, their environmental impact and their peoples happiness. 'Happy Life Years' is the amount of years a person can expect to be enjoyable and less stressful in the given country. 'Inequality adjusted life expectancy' is life expectancy but is when you take in to account the different living standards of different groups, different incomes, different ethnicity and others. The data set was compiled using many open polls and data sets from the UN, York University and others.

We have analysed our data through three categories: geographic, economic and mental health. The first category contains variables population, ecological footprint and location (continent). Our second category compares values using variables average well-being,happy life years,inequality of outcomes, inequality-adjusted life expectancy, and inequality-adjusted well-being. 

We will examine the changes within these three categories through years 2015 to 2020 and compare specifically years 2015-2019 and 2020 to understand the effects of the pandemic on the countries' happy planet indexes.



Write-up of your project and findings go here. 
Think of this as the text of your presentation. 
The length should be roughly 5 minutes when read out loud.
Although pacing varies, a 5-minute speech is roughly 750 words. 
To use the word count addin, select the text you want to count the words of (probably this is the 
Summary section of this document, go to Addins, and select the `Word count` addin).
This addin counts words using two different algorithms, but the results should be similar and as long as you're in the ballpark of 750 words, you're good! 
The addin will ignore code chunks and only count the words in prose.

You can also load your data here and present any analysis results / plots, but I strongly urge you to keep that to a minimum (maybe only the most important graphic, if you have one you can choose). 
And make sure to hide your code with  `echo = FALSE` unless the point you are trying to make is about the code itself.
Your results with proper output and graphics go in your presentation, this space is for a brief summary of your project.

```{r load-data, echo = FALSE}
library(tidyverse)
library(janitor)
library(readxl)
hpi_2016 <- read_excel("hpi_2016_data.xlsx") %>% 
  clean_names()
hpi_2017 <- read_excel("hpi_2017_data.xlsx") %>% 
  clean_names()
hpi_2018 <- read_excel("hpi_2018_data.xlsx") %>% 
  clean_names()
hpi_2019 <- read_excel("hpi_2019_data.xlsx") %>% 
  clean_names()
hpi_2020 <- read_excel("hpi_2020_data.xlsx") %>% 
  clean_names()
```

```{r cleaning-data, echo = FALSE}
hpi_2016 <- hpi_2016 %>%
  mutate(year = 2016)
hpi_2017 <- hpi_2017 %>%
  mutate(year = 2017)
hpi_2018 <- hpi_2018 %>%
  mutate(year = 2018)
hpi_2019 <- hpi_2019 %>%
  mutate(year = 2019)
hpi_2020 <- hpi_2020 %>%
  mutate(year = 2020)

#The idea to use bind_rows came from the following post on StackOverflow: https://stackoverflow.com/questions/48444790/tidydata-combining-two-dataset-from-different-years-r-to-make-a-tidy-data-se
hpi_data <- bind_rows(hpi_2016, hpi_2017)
hpi_data <- bind_rows(hpi_data, hpi_2018)
hpi_data <- bind_rows(hpi_data, hpi_2019)
hpi_data <- bind_rows(hpi_data, hpi_2020)

hpi_data <- hpi_data %>%
  rename("pop_1000s"     = population_thousands,
         "life_exp"      = life_expectancy_years,
         "wellbeing"     = ladder_of_life_wellbeing_0_10,
         "eco_footprint" = ecological_footprint_g_ha,
         "biocapacity"   = biocapacity_for_year_g_ha,
         "gdp"           = gdp_per_capita,
         "region"        = continent) %>%
  transform(gdp = as.numeric(gdp))
  
```

```{r glimpse data}
glimpse(hpi_data)
```

## Presentation

Our presentation can be found [here](presentation/presentation.html).

## Data 

Include a citation for your data here. 
See http://libraryguides.vu.edu.au/c.php?g=386501&p=4347840 for guidance on proper citation for datasets. 
If you got your data off the web, make sure to note the retrieval date.

## References

List any references here. You should, at a minimum, list your data source.

