---
title: "Project proposal"
author: "MZM"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

*For instructions on what each section should include, please see the [project page](https://idsed.digital/assessments/project/#proposal) on the course website. Remove this text when completing your proposal*.

## 1. Introduction

In our project we aim to show how different countries compare against one another in multiple variables, one of which being the 'Happy Planet Index'. This will help us understand which regions or countries are able to provide a good living standard while also having a healthy environmental footprint. Our data is from a data set named the 'Happy Planet Index' which has taken most countries and used their social happiness, environmental impact and also inequalities and differences of the countries people. Some important variables to know are the 'Happy Planet Index' which is a combination of a countries inequalities of living, their environmental impact and their peoples happiness. 'Happy Life Years' is the amount of years a person can expect to be enjoyable and less stressful in the given country. 'Inequality adjusted life expectancy' is life expectancy but is when you take in to account the different living standards of different groups, different incomes, different ethnicity and others. The data set was compiled using many open polls and data sets from the UN, York University and others.

## 2. Data

```{r load-data, echo = FALSE}
library(tidyverse)
library(janitor)
library(readxl)
hpi_2016 <- read_excel("hpi_2016_data.xlsx") %>% 
  clean_names()
hpi_2017 <- read_excel("hpi_2017_data.xlsx") %>% 
  clean_names()
hpi_2018 <- read_excel("hpi_2018_data.xlsx") %>% 
  clean_names()
hpi_2019 <- read_excel("hpi_2019_data.xlsx") %>% 
  clean_names()
hpi_2020 <- read_excel("hpi_2020_data.xlsx") %>% 
  clean_names()

```

## 3. Data analysis plan

We have analysed our data through three categories: geographic, economic and mental health. The first category contains variables population, ecological footprint and location (continent). Our second category compares values using variables average well-being,happy life years,inequality of outcomes, inequality-adjusted life expectancy, and inequality-adjusted well-being. 

We will examine the changes within these three categories through years 2016 to 2020 and compare specifically years 2016-2019 and 2020 to understand the effects of the pandemic on the countries' happy planet indexes.

Method: We are planning on taking the average of values of data tables from years 2016 to 2019 and creating another table. Comparing the new data frame, containing the averages of all values for the necessary variables, with values of 2020 will give us a more reliable difference/change due to the pandemic. Our analysis will contain three graphs (one for each category) with all the necessary variables.

Conclusion: We expect decrease in mental health and economic factors due to the pandemic.
