---
title: "Project proposal"
author: "MZM"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readxl)
library(janitor)
library(scales)
```


## 1. Introduction

In our project we aim to show how different countries compare against one another in multiple variables, one of which being the 'Happy Planet Index'. This will help us understand which regions or countries are able to provide a good living standard while also having a healthy environmental footprint. Our data is from a data set named the 'Happy Planet Index' which has taken most countries and used their social happiness, environmental impact and also inequalities and differences of the countries people. Some important variables to know are the 'Happy Planet Index' which is a combination of a countries inequalities of living, their environmental impact and their peoples happiness. 'Eco Footprint' is the average demand for the natural resources of the environment, so how much land and resources would be needed to sustain the country. 'Biocapacity' is the average wellbeing reported by the citizens of a country, measured through a survey in which people ranked their wellbeing from 1 to 10. The data set was compiled using many open polls and data sets from the UN, York University and others.

## 2. Data

```{r load-data, echo = FALSE}
hpi_2016 <- read_excel("/cloud/project/data/hpi_2016_data.xlsx") %>% 
  clean_names()
hpi_2017 <- read_excel("/cloud/project/data/hpi_2017_data.xlsx") %>% 
  clean_names()
hpi_2018 <- read_excel("/cloud/project/data/hpi_2018_data.xlsx") %>% 
  clean_names()
hpi_2019 <- read_excel("/cloud/project/data/hpi_2019_data.xlsx") %>% 
  clean_names()
hpi_2020 <- read_excel("/cloud/project/data/hpi_2020_data.xlsx") %>% 
  clean_names()

```

```{r cleaning-data, echo = FALSE}
hpi_2016 <- hpi_2016 %>%
  mutate(year = 2016)
hpi_2017 <- hpi_2017 %>%
  mutate(year = 2017)
hpi_2018 <- hpi_2018 %>%
  mutate(year = 2018)
hpi_2019 <- hpi_2019 %>%
  mutate(year = 2019)
hpi_2020 <- hpi_2020 %>%
  mutate(year = 2020)

#The idea to use bind_rows came from the following post on StackOverflow: https://stackoverflow.com/questions/48444790/tidydata-combining-two-dataset-from-different-years-r-to-make-a-tidy-data-se
hpi_data <- bind_rows(hpi_2016, hpi_2017)
hpi_data <- bind_rows(hpi_data, hpi_2018)
hpi_data <- bind_rows(hpi_data, hpi_2019)
hpi_data <- bind_rows(hpi_data, hpi_2020)

hpi_data <- hpi_data %>%
  rename("pop_1000s"     = population_thousands,
         "life_exp"      = life_expectancy_years,
         "wellbeing"     = ladder_of_life_wellbeing_0_10,
         "eco_footprint" = ecological_footprint_g_ha,
         "biocapacity"   = biocapacity_for_year_g_ha,
         "gdp"           = gdp_per_capita,
         "region"        = continent) %>%
  transform(gdp = as.numeric(gdp))

hpi_data <- hpi_data %>%
  mutate(region = case_when(
    region == 1 ~ "Latin America",
    region == 2 ~ "North America & Oceania",
    region == 3 ~ "Western Europe",
    region == 4 ~ "Middle East",
    region == 5 ~ "Africa",
    region == 6 ~ "South Asia",
    region == 7 ~ "Eastern Europe & Central Asia",
    region == 8 ~ "East Asia"
  ))
```

```{r glimpse data}
glimpse(hpi_data)
```
## 3. Data analysis plan
```{r wellbeing-time plot}
hpi_data %>%
  ggplot(aes(x = year, y = wellbeing, group = year)) +
  geom_violin(trim = FALSE) +
  facet_wrap(~region) +
  labs(
    title = "Wellbeing around the world",
    subtitle = "Over time",
    x = "Years",
    y = "Wellbeing measure"
  )

hpi_data %>%
  ggplot(aes(x = eco_footprint, y = wellbeing, group = year, color= region)) +
  geom_line(trim = FALSE) +
  facet_wrap() +
  labs(
    title = "Wellbeing around the world",
    subtitle = "Over time",
    x = "Ecological Footprint",
    y = "Wellbeing measure"
  )
hpi_data %>%
  ggplot(aes(x = life_exp, y = wellbeing, group = year)) +
  geom_point(trim = FALSE) +
  facet_wrap(~region) +
  labs(
    title = "Wellbeing around the world",
    subtitle = "Over time",
    x = "Life Expectancy",
    y = "Wellbeing measure"
  )

library(viridis)
library(gganimate)

hpi_data_clean <- na.omit(hpi_data)

p <- ggplot(
  hpi_data_clean, 
  aes(x = wellbeing, y=life_exp, size = pop_1000s, colour = country)
  ) +
  geom_point(show.legend = FALSE, alpha = 0.7) +
  scale_color_viridis(name= "country", discrete = TRUE) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  guides(colour = guide_legend(order = 2), shape = guide_legend(order = 1)) +
  theme_void() + 
    theme(legend.position = "bottom", legend.box = "vertical", legend.title.align = 0) +
    theme(plot.title = element_text(hjust = 0.5, vjust = 0.05)) +
    theme(plot.caption = element_text(hjust = 0, color="gray40", size=10)) +
  labs(x = "Wellbeing", y = "Life Expectancy")

animation<- p + transition_states(year,
                              transition_length = 2,
                              state_length = 1) 
gifski_renderer(file = NULL, loop = TRUE, width = NULL, height = NULL)(animation)
animation
library(reprex)

file_renderer(dir = ".", prefix = "gganim_plot", overwrite = FALSE)(animation)
```
It is clear from the graph that people in north america and oceania are the happiest about their wellbeing standards and that south asia seems very varied but also the least content. There doesn't seem to be a clear impact from COVID in the wellbeing of countries.

```{r gdp-life_exp plot}
hpi_data %>%
  ggplot(aes(x = gdp, y = life_exp, colour = region)) +
  geom_point() +
  facet_wrap(~year) +
  labs(
    title = "GDP and life expectancy around the world",
    subtitle = "Over time",
    x = "GDP per capaita (in thousands)",
    y = "Life expectancy",
    region = "Region"
  ) +
  theme(legend.position = "bottom") +
  scale_x_continuous(labels = label_dollar(scale = 0.001,
                                           prefix = "$",
                                           suffix = "K"))
```
As we can see there is a clear relationship between life expectancy and GDP, countries with higher life expectancy also tend to have higher GDP's. Most countries with low GDP's and low life expectancy are in africa meanwhile western europe holds many of the top spots. There isnt much change year to year.
```{r}
hpi_data %>%
  ggplot(aes(x = eco_footprint, fill = region)) +
  geom_histogram() +
  facet_wrap(~region) +
  labs(
    title = "Ecological footprint of countries",
    subtitle = "Faceted by region",
    x = "Ecological footprint (in global hectares)",
    y = "Number of countries"
  ) +
  theme(legend.position = "none")

```
The graphs show the ecological footprint of each country 5 times for each year. We can see that Africa has a very low ecological footprint in general likely due to low rates of industrialisation. The middle east has the highest ecological footprint countries and the lowest simultaneously, this could be due to some countries being heavily industrialised and others having vast deserts with no ecological footprint.

We have analysed our data through three categories: geographic, economic and mental health. The first category contains variables on general well being of the population. our second category obtainslife expactancyand GDP per capita. Lastly, our third category has information on ecological footprint based on their location (continent).

We will examine the changes within these three categories through years 2016 to 2020 and compare specifically years 2016-2019 and 2020 to understand the effects of the pandemic on the countries' happy planet indexes.

Method: We are planning on taking the average of values of data tables from years 2016 to 2019 and creating another table. Comparing the new data frame, containing the averages of all values for the necessary variables, with values of 2020 will give us a more reliable difference/change due to the pandemic. Our analysis will contain three graphs (one for each category) with all the necessary variables.

Conclusion: We expect decrease in mental health and economic factors due to the pandemic.
