---
title: "Project_Research"
author: "MZM"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, echo=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(scales)
library(splines)
library(modelr)
library(tidymodels)
```

```{r functions, echo=FALSE}
tidy_name_year <- function(data, year_no){
  data <- data %>%
    clean_names() %>%
    mutate(year = year_no)
  return(data)
}

tidy_append_data <- function(data1, data2, data3, 
                             data4, data5){
  data <- bind_rows(data1, data2)
  data <- bind_rows(data, data3)
  data <- bind_rows(data, data4)
  data <- bind_rows(data, data5)
  
  data <- data %>%
  rename("pop_1000s"     = population_thousands,
         "life_exp"      = life_expectancy_years,
         "wellbeing"     = ladder_of_life_wellbeing_0_10,
         "eco_footprint" = ecological_footprint_g_ha,
         "biocapacity"   = biocapacity_for_year_g_ha,
         "gdp"           = gdp_per_capita,
         "region"        = continent) %>%
  transform(gdp = as.numeric(gdp))
  
  data <- data %>%
  mutate(region = case_when(
    region == 1 ~ "Latin America",
    region == 2 ~ "North America & Oceania",
    region == 3 ~ "Western Europe",
    region == 4 ~ "Middle East",
    region == 5 ~ "Africa",
    region == 6 ~ "South Asia",
    region == 7 ~ "Eastern Europe & Central Asia",
    region == 8 ~ "East Asia"),
    happy_life_years = life_exp * wellbeing / 10)
}
```

```{r load data, echo=FALSE}
hpi_2016 <- read_excel("/cloud/project/data/hpi_2016_data.xlsx")
hpi_2017 <- read_excel("/cloud/project/data/hpi_2017_data.xlsx")
hpi_2018 <- read_excel("/cloud/project/data/hpi_2018_data.xlsx") 
hpi_2019 <- read_excel("/cloud/project/data/hpi_2019_data.xlsx")
hpi_2020 <- read_excel("/cloud/project/data/hpi_2020_data.xlsx")

hpi_2016 <- tidy_name_year(hpi_2016, 2016)
hpi_2017 <- tidy_name_year(hpi_2017, 2017)
hpi_2018 <- tidy_name_year(hpi_2018, 2018)
hpi_2019 <- tidy_name_year(hpi_2019, 2019)
hpi_2020 <- tidy_name_year(hpi_2020, 2020)

hpi_data <- tidy_append_data(hpi_2016, hpi_2017, hpi_2018, 
                             hpi_2019, hpi_2020)
```

```{r gdp-life_exp plot}
hpi_data %>%
  ggplot(aes(x = gdp, y = life_exp, colour = region)) +
  geom_point() +
  facet_wrap(~year) +
  labs(
    title = "GDP and life expectancy around the world",
    subtitle = "Over time",
    x = "GDP per capaita (in thousands)",
    y = "Life expectancy",
    region = "Region"
  ) +
  theme(legend.position = "bottom") +
  scale_x_continuous(labels = label_dollar(scale = 0.001,
                                           prefix = "$",
                                           suffix = "K"))
```

```{r gdp-hpi plot}
hpi_data %>%
  filter(year != 2020) %>%
  ggplot(aes(x = gdp, y = hpi, colour = region)) +
  geom_point(alpha = 0.3)
```

```{r happy life years and gdp plot}
hpi_data %>%
  filter(year != 2020) %>%
  ggplot(aes(x = gdp, y = happy_life_years, color = region)) +   geom_point() +
  facet_wrap(~year)
```

```{r happy life years and gdp plots}
summarised_gdp_hpi <- hpi_data %>%
  filter(year != 2020) %>%
  group_by(country, region) %>%
  summarise(avg_gdp = median(gdp), 
            avg_life_years = mean(happy_life_years, 
                                  rm.na = TRUE))

summarised_gdp_hpi %>%
  ggplot(aes(x = avg_gdp, y = avg_life_years, 
             colour = region)) +
  geom_point() +
  geom_smooth(method = "lm", 
              formula = y~x,
              se =  FALSE,
              colour = "black") +
  facet_wrap(~region, nrow = 2) +
  theme(legend.position = "bottom")
```

```{r lm for South Asia}
gdp_hpi_SA <- summarised_gdp_hpi %>%
  filter(region == "South Asia")

linear_reg() %>%
  set_engine("lm") %>%
  fit(avg_life_years ~ avg_gdp, data = gdp_hpi_SA) %>%
  tidy()
```

```{r lm for Eastern Europe & Central Asia}
gdp_hpi_EECA <- summarised_gdp_hpi %>%
  filter(region == "Eastern Europe & Central Asia")

linear_reg() %>%
  set_engine("lm") %>%
  fit(avg_life_years ~ avg_gdp, data = gdp_hpi_EECA) %>%
  tidy()
```

```{r lm for Middle East}
gdp_hpi_ME <- summarised_gdp_hpi %>%
  filter(region == "Middle East")

linear_reg() %>%
  set_engine("lm") %>%
  fit(avg_life_years ~ avg_gdp, data = gdp_hpi_ME) %>%
  tidy()
```

```{r lm for Latin America}
gdp_hpi_LA <- summarised_gdp_hpi %>%
  filter(region == "Latin America")

linear_reg() %>%
  set_engine("lm") %>%
  fit(avg_life_years ~ avg_gdp, data = gdp_hpi_LA) %>%
  tidy()
```

```{r lm for North America & Oceania}
gdp_hpi_NAO <- summarised_gdp_hpi %>%
  filter(region == "North America & Oceania")

linear_reg() %>%
  set_engine("lm") %>%
  fit(avg_life_years ~ avg_gdp, data = gdp_hpi_NAO) %>%
  tidy()
```

```{r lm for Western Europe}
gdp_hpi_WE <- summarised_gdp_hpi %>%
  filter(region == "Western Europe")

linear_reg() %>%
  set_engine("lm") %>%
  fit(avg_life_years ~ avg_gdp, data = gdp_hpi_WE) %>%
  tidy()
```

```{r lm for Africa}
gdp_hpi_A <- summarised_gdp_hpi %>%
  filter(region == "Africa")

linear_reg() %>%
  set_engine("lm") %>%
  fit(avg_life_years ~ avg_gdp, data = gdp_hpi_A) %>%
  tidy()
```

```{r lm for East Asia}
gdp_hpi_EA <- summarised_gdp_hpi %>%
  filter(region == "East Asia")

linear_reg() %>%
  set_engine("lm") %>%
  fit(avg_life_years ~ avg_gdp, data = gdp_hpi_EA) %>%
  tidy()
```

```{r hpi years plot}
hpi_data %>%
  group_by(country, )
  ggplot(aes(x = country, y = hpi)) +
  geom_point() +
  facet_grid(~year)
```

```{r eco-analysis-middle-east}
hpi_3 <- filter(hpi_data, region == c("Africa", "Middle East", "Western Europe"))
ggplot(data = hpi_3, aes(x = year, y = eco_footprint, group = country))+
  geom_line()+
  facet_wrap(~region, nrow = 1)
```


```{r eco-analysis-western-europe}
hpi_western_europe <- filter(hpi_data, region == "Western Europe", eco_footprint > 5, eco_footprint < 9)

ggplot(data = hpi_western_europe, aes(x = year, y = eco_footprint, colour = country))+
  geom_line()
```


```{r eco-analysis-africa}
  ggplot(data = hpi_data, aes(x = year, y = eco_footprint, group = country))+
  geom_line()
```

```{r eco-gdp}
  ggplot(data = hpi_data, aes(x = gdp, y = eco_footprint, group = country, colour = gdp))+
  geom_jitter()
```

```{r eco-gdp}
target = c("Middle East", "Africa", "Western Europe")

gdp_3 <- filter(hpi_data, region == target, year == 2019)

ggplot(data = gdp_3, aes(x = country, y = gdp, fill = region, group = region))+
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90))
```

From 2016 to 2019 most countries have an increasing or steady ecological footprint but from
2019 to 2020 we see a decline from virtually every nation observed. 
This is not only a trend followed by similar countries as the gdp of countries in
these regions vary significantly with africa being a lot lower than the rest. There are 
many possible reasons for this change but most obvious would be covid.
The large scale closures and lockdowns would have led to less resources being required
thus lowering countries ecological footprint. Other reasons could be governments being
more proactive about climate change but this would not explain how wide spread this
decrease is.